### 2016-06-22 paper/icfp16-cameraready/gallery2.tex
actually, let's group all of the graphs
leave the traces inline

\subsection{Witness Utility}\label{sec:advantage-traces}

@@

\paragraph{Example: Recursion with Bad Operator}
The recursive function @sqsum@ should square each
element of the input list and then compute the sum
of the result.
%
\begin{ecode}
  let rec sqsum xs = match xs with
    | [] -> 0
    | h::t -> __(sqsum t)__ @ (h * h)
\end{ecode}
%
Unfortunately the student has used the list-append
operator |@| instead of \texttt{+} to compute the sum.
%
Both \ocaml\ and \sherrloc\ blame the \emph{wrong location},
namely the recursive call @sqsum t@ with the message
%
\begin{verbatim}
  This expression has type
    int
  but an expression was expected of type
    'a list
\end{verbatim}
%
\toolname\ produces the following trace showing how the evaluation of
@sqsum [1]@ gets stuck:
%
\begin{center}
  \includegraphics[height=125px]{sqsum.png}
\end{center}
%
The figure highlights the entire stuck term
(not just the recursive call), emphasizing
the \emph{conflict} between @int@ and @list@
rather than assuming one or the other is correct.

\paragraph{Example: Recursion with Bad Base Case}
%
The function @sumList@ should add up
the elements of its input list.
%
\begin{ecode}
  let rec sumList xs = match xs with
    | []    -> ==[]==
    | y::ys -> y + __sumList ys__
\end{ecode}
%
Unfortunately, in the base case, it returns @[]@
instead of @0@.
%
\sherrloc\ blames the base case, and \ocaml\
assumes the base case is correct and blames
the \emph{recursive call} on line 3:
%
\begin{verbatim}
  This expression has type
    'a list
  but an expression was expected of type
    int
\end{verbatim}
%
Both of the above are parts of the full story, which
is summarized by \toolname's trace showing
how @sumList [1; 2]@ gets stuck at @2 + []@.
%
\begin{center}
  \includegraphics[height=125px]{sumlist.png}
\end{center}
%
The trace clarifies immediately (via the third step)
that the @[]@ is the result of the recursive call
@sumList []@, and shows how it is incompatible with
the subsequent \texttt{+} operation.

@@

\pagebreak
\paragraph{Example: Bad Helper Function that Type-Checks}
%
The function @digitsOfInt@ is supposed to return a list of
the digits of the input integer.
%
\begin{ecode}
  let append x xs =
    match xs with
    | [] -> [x]
    | _  -> x :: xs

  let rec digitsOfInt n =
    if n <= 0 then
      []
    else
      append (==digitsOfInt (n / 10)==)
             [__n mod 10__]
\end{ecode}
%
Unfortunately, the student's @append@ function \emph{conses} an element
onto a list instead of appending two lists.
%
Though incorrect, @append@ still type-checks and thus \ocaml and
\sherrloc blame the \emph{use-site} on line 10.
%
\begin{verbatim}
  This expression has type
    int
  but an expression was expected of type
    'a list
\end{verbatim}
%
\toolname, in contrast, makes no assumptions about @append@ and produces
a trace that illustrates the true error on line 4, by
highlighting the conflict in consing a list onto a list of integers.
%
\begin{center}
  \includegraphics[height=175px]{digitsOfInt.png}
\end{center}
%

>>>>

\subsubsection{Qualitative}\label{sec:advantage-traces}

@@

% \begin{figure*}[htp]
%   \centering
%   \includegraphics[height=125px]{sqsum.png}
%   \includegraphics[height=125px]{sumlist.png}
%   \includegraphics[height=150px]{digitsOfInt.png}
%   \includegraphics[height=125px]{wwhile.png}
%   \caption{(Left to right) Jump-compressed traces showing how
%     \texttt{sqsum}, \texttt{sumList}, \texttt{digitsOfInt}, and
%     \texttt{wwhile} go wrong in \S~\ref{sec:advantage-traces}.}
%   \label{fig:traces}
% \end{figure*}

\paragraph{Example: Recursion with Bad Operator}
The recursive function @sqsum@ should square each
element of the input list and then compute the sum
of the result.
%
\begin{ecode}
  let rec sqsum xs = match xs with
    | [] -> 0
    | h::t -> __(sqsum t)__ @ (h * h)
\end{ecode}
%
Unfortunately the student has used the list-append
operator |@| instead of \texttt{+} to compute the sum.
%
Both \ocaml\ and \sherrloc\ blame the \emph{wrong location},
namely the recursive call @sqsum t@ with the message
%
\begin{verbatim}
  This expression has type
    int
  but an expression was expected of type
    'a list
\end{verbatim}
%
\toolname\ produces a trace showing how the evaluation of
@sqsum [1]@ gets stuck.
%
\begin{center}
  \includegraphics[height=125px]{sqsum.png}
\end{center}
%
The trace highlights the entire stuck term
(not just the recursive call), emphasizing
the \emph{conflict} between @int@ and @list@
rather than assuming one or the other is correct.

\paragraph{Example: Recursion with Bad Base Case}
%
The function @sumList@ should add up
the elements of its input list.
%
\begin{ecode}
  let rec sumList xs = match xs with
    | []    -> ==[]==
    | y::ys -> y + __sumList ys__
\end{ecode}
%
Unfortunately, in the base case, it returns @[]@
instead of @0@.
%
\sherrloc\ blames the base case, and \ocaml\
assumes the base case is correct and blames
the \emph{recursive call} on line 3:
%
\begin{verbatim}
  This expression has type
    'a list
  but an expression was expected of type
    int
\end{verbatim}
%
Both of the above are parts of the full story, which
is summarized by \toolname's trace showing
how @sumList [1; 2]@ gets stuck at @2 + []@.
%
\begin{center}
  \includegraphics[height=125px]{sumlist.png}
\end{center}
%
The trace clarifies (via the third step)
that the @[]@ is the result of the recursive call
@sumList []@, and shows how it is incompatible with
the subsequent \texttt{+} operation.

@@

%\pagebreak
\paragraph{Example: Bad Helper Function that Type-Checks}
%
The function @digitsOfInt@ is supposed to return a list of
the digits of the input integer.
%
\begin{ecode}
  let append x xs =
    match xs with
    | [] -> [x]
    | _  -> x :: xs

  let rec digitsOfInt n =
    if n <= 0 then
      []
    else
      append (==digitsOfInt (n / 10)==)
             [__n mod 10__]
\end{ecode}
%
%\pagebreak
Unfortunately, the student's @append@ function \emph{conses} an element
onto a list instead of appending two lists.
%
Though incorrect, @append@ still type-checks and thus \ocaml and
\sherrloc blame the \emph{use-site} on line 10.
%
\begin{verbatim}
  This expression has type
    int
  but an expression was expected of type
    'a list
\end{verbatim}
%
\toolname, in contrast, makes no assumptions about @append@ and produces
a trace that illustrates the true error on line 4, by
highlighting the conflict in consing a list onto a list of integers.
%
\begin{center}
  \includegraphics[height=175px]{digitsOfInt.png}
\end{center}
%
